Bayesian Exploration in Deep Reinforcement Learning

### 2022-02-10 papers/main.tex
TEMPORARY DEPLOY COMMIT

% \documentclass{uai2021} % for initial submission
\documentclass[]{uai2021} % after acceptance, for a revised
                                    % version; also before submission to
                                    % see how the non-anonymous paper
                                    % would look like
%% There is a class option to choose the math font
% \documentclass[mathfont=ptmx]{uai2021} % ptmx math instead of Computer
                                         % Modern (has noticable issues)
% \documentclass[mathfont=newtx]{uai2021} % newtx fonts (improves upon
                                          % ptmx; less tested, no support)
% NOTE: Only keep *one* line above as appropriate, as it will be replaced
%       automatically for papers to be published. Do not make any other
%       change above this note for an accepted version.

@@

@@

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods has also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

@@

\citet{fortunato_noisy_2019} introduced NoisyNet for exploration.
These are networks with stochastic weights, where each weight \(w^{(l)}_{ij}\)
has an added pertrubaition sampled from a noise distribution with standard deviation \(\sigma^{(l)}_{ij}\).
After initializing \(\bm{w}\) and \(\bm{\sigma}\) such that the network has sufficient
stochasticity for exploration, both parameters are learned using standard backpropagation.
The approach is similar to variational inference schemes such as Bayes by backprop~\citep{blundell_weight_2015} where the weights of a neural network model are assumed to be normally distributed with mean 
\(\mu^{(l)}_{ij}\) and standard deviation \(\sigma^{(l)}_{ij}\). The objective of
Bayes by backprop, however, is to approximate the posterior distribution
\(p(\bm{w} \vert \mathcal{D})\) in a task with a fixed dataset \(\mathcal{D}\).
The parameter distribution in NoisyNet does not necessarily
converge to an approximate posterior. This means that it does not have the same
guarantee on total regret as methods that approximate a posterior over the
value functions~\citep{osband_generalization_2016}. They do, however, have
experimental results which shows that the function does not always converge to
a deterministic solution, but it is unclear why the network learns to introduce
more noise into the network parameters.

@@

\subsection{Functional Variational Bayesian Neural Networks}
There are several ways of approximating the posterior distribution of the weights
in a neural network with respect to a prior and a dataset. Typically, the dataset
\(\D\) is static and with datapoints \(\bm{x}\) and labels \(\bm{y}\), and the 
prior is defined as a distribution over the weights
\(p(\bm{w})\)~\citep{rezende_stochastic_2014,blundell_weight_2015,ritter_scalable_2018,maddox_simple_2019}.
By defining a prior over the weights, they can use approximate variational inference
methods to approximate \(p(\bm{w} \vert \D) = \frac{p(\D \vert \bm{w})p(\bm{w})}{p(\D)}\).
The disadvantage of this is that \(p(\bm{w})\) only acts as a regularizer, and is not
used to incorporate prior knowledge. Any prior knowledge we might have about the optimizer
function would be very difficult to translate into a prior distribution in the weight space.

Functional variational Bayesian neural networks \citep{sun_functional_2019} is 
a variational inference method for neural networks that approximates the 
posterior distribution in function space. This means that our prior will be a distribution
over functions, also known as a stochastic process.
\citet{sun_functional_2019} show that for two stochastic processes \(A\) and \(B\), the
KL-divergence from \(A\) to \(N\) is the supremum of the marginal KL-divergences over
all finite measurement sets. Let \(A_\X\) be the marginal distribution of function values
at \(\X \in \mathcal{X}^n\), then:

@@

We will use this identity to approximate the functional KL-divergence by sampling 
measurement sets \(\X \in \mathcal{X}^n\). Note that because this KL-divergence is a 
supremum, sampling measurment sets will always give us an underestimate of the true
KL-divergence.

In our case \(\mathcal{X} = \state \times \action\). First we start by
defining \(\X^D\) as the set of datapoints \(\D \in \state \times \action\) that we have
explored, but we also need points outside of what we have already explored. Sampling points
from undiscovered states is difficult, however, since we typically don't know what the set
of possible states includes. Since we know that each state only supports a finite number of
actions, we propose to let \(\X^M\) be a set of state-action pairs that 
The measurement set \(\X\) is created by concatenating the training points \(\X^D\) and a
set of \(M\) points \(\X^M\) drawn from some distribution \(c\) that has full support
in the space of interesting test cases.

@@

The first term is just the expected value of the score function, which is zero.
The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log \rho_\phi(\f(\X))\)
and \(\nabla_\f \log p(\f(\X))\). \(\nabla_\f \log \rho_\phi(\f(\X))\) is likely 
intractable, considering \(\rho_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f(\X))\)
can be easy to compute analytically. To reduce variance in the gradients, we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log \rho_\phi(\f(\X))\), they use the Spectral Stein Gradient Estimator (SSGE)
\citep{shi_spectral_2018}. In short, this is a method for estimating the gradient
function of implicit distributions using approximations to eigenfunctions of a 
kernel-based operator. We bring three hyperparameters into the algorithm from this
method, the number of samples \(k\) from the implicit distribution used to approximate
the gradient, the number of eigenvectors \(J\) used to approximate the gradient, and
\(\eta\), a regularization parameter that smooths the gradient function.

@@

\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge. The training curves
of our model compared to that of standard NoisyNet-DQN on the four selected OpenAI
Gym~\citep{brockman_openai_2016} environments is shown in Figure~\ref{fig:training_curves}.
We hypothesize that the added KL-term will push the algorithm to retain weight uncertainty high
in order to keep a higher uncertainty in the output. While both methods can
find an optimal policy in Cartpole, our method uses considerably fewer frames
to do so.

@@

\cite{fortunato_noisy_2019} noted that the learned variance in their weight increased
in some games despite there existing an optimal deterministic solution, and the loss
provides no incentive to maintain any uncertainty. Figure~\ref{fig:mean_std} shows
the mean-absolute standard deviation for the penultimate and final layer. It is 
interesting to see that the standard deviation for NoisyNet continues to decrease
throughout the training while LFM decreases faster, but then flattens out earlier.
This seems to indicate that it has found a stable policy, where optimizing further would
be overfitting to noise.

@@

\subsection{Informative Prior}
One of the benefits of a functional prior is that we can incorporate domain knowledge to
get more efficient exploration. We will now look at how our method can utilize domain
knowledge to improve sample efficiency. To measure the effectiveness of priors we will use
a distribution that has a higher concentration of large Q-values for actions that we want
to incentivize.

>>>>

% \documentclass{uai2021} % for initial submission
\documentclass[]{uai2022} % after acceptance, for a revised
                                    % version; also before submission to
                                    % see how the non-anonymous paper
                                    % would look like
%% There is a class option to choose the math font
% \documentclass[mathfont=ptmx]{uai2021} % ptmx math instead of Computer
                                         % Modern (has noticable issues)
% \documentclass[mathfont=newtx]{uai2021} % newtx fonts (improves upon
                                          % ptmx; less tested, no support)
% NOTE: Only keep *one* line above as appropriate, as it will be replaced
%       automatically for papers to be published. Do not make any other
%       change above this note for an accepted version.

@@

% \setlength{\tabcolsep}{12pt}

@@

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
A learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy, is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution and selecting the action that gave the highest
reward. Bayesian methods have also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

@@

\citet{fortunato_noisy_2019} introduced NoisyNet for exploration.
These are networks with stochastic weights, where each weight \(w^{(l)}_{ij}\)
has an added perturbation sampled from a noise distribution with standard deviation \(\sigma^{(l)}_{ij}\).
After initializing \(\bm{w}\) and \(\bm{\sigma}\) such that the network has sufficient
stochasticity for exploration, both parameters are learned using standard backpropagation.
The approach is similar to variational inference schemes such as Bayes by backprop~\citep{blundell_weight_2015} where the weights of a neural network model are assumed to be normally distributed with mean 
\(\mu^{(l)}_{ij}\) and standard deviation \(\sigma^{(l)}_{ij}\). The objective of
Bayes by backprop, however, is to approximate the posterior distribution
\(p(\bm{w} \vert \mathcal{D})\) in a task with a fixed dataset \(\mathcal{D}\).
The parameter distribution in NoisyNet does not necessarily
converge to an approximate posterior. This means that it does not have the same
guarantee on total regret as methods that approximate a posterior over the
value functions~\citep{osband_generalization_2016}. They do, however, have
experimental results which show that the function does not always converge to
a deterministic solution. It is unclear why the network learns to introduce
more noise into the network parameters.

@@

\subsection{Functional Variational Bayesian Neural Networks}
There are several ways of approximating the posterior distribution of the weights
in a neural network with respect to a prior and a dataset. Typically, the dataset
\(\D\) is given and with datapoints \(\bm{x}\) and labels \(\bm{y}\), and the 
prior is defined as a distribution over the weights
\(p(\bm{w})\)~\citep{rezende_stochastic_2014,blundell_weight_2015,ritter_scalable_2018,maddox_simple_2019}.
By defining a prior over the weights, they can use approximate variational inference
methods to approximate \(p(\bm{w} \vert \D) = \frac{p(\D \vert \bm{w})p(\bm{w})}{p(\D)}\).
The disadvantage of this is that \(p(\bm{w})\) only acts as a regularizer, and is not
used to incorporate prior knowledge. Any prior knowledge we might have about the target
function would be very difficult to translate into a prior distribution in the weight space.

Functional variational Bayesian neural networks \citep{sun_functional_2019} is 
a variational inference method for neural networks that approximates the 
posterior distribution in function space. This means that our prior will be a distribution
over functions, also known as a stochastic process.
\citet{sun_functional_2019} show that for two stochastic processes \(A\) and \(B\), the
KL-divergence from \(A\) to \(B\) is the supremum of the marginal KL-divergences over
all finite measurement sets. Let \(A_\X\) be the marginal distribution of function values
at \(\X \in \mathcal{X}^n\), then:

@@

We will use this identity to approximate the functional KL-divergence by sampling 
measurement sets \(\X \in \mathcal{X}^n\). Note that because this KL-divergence is a 
supremum, sampling measurement sets will always give us an underestimate of the true
KL-divergence.

In our case \(\mathcal{X} = \state \times \action\) and \(\X = \begin{bmatrix} s^T & a^T \end{bmatrix}^T\). First we start by
defining \(\X^D\) as the set of data points \(\D \in \state \times \action\) that we have
explored, but we also need points outside what we have already explored. Sampling points
from undiscovered states is difficult, because we typically do not know what the set
of possible states includes. Since we know that each state only supports a finite number of
actions, we propose to let \(\X^M\) be a set of state-action pairs that 
The measurement set \(\X\) is created by concatenating the training points \(\X^D\) and a
set of \(M\) points \(\X^M\) drawn from some distribution \(c\) that has full support
in the space of interesting test cases.

@@

The first term is just the expected value of the score function, which is zero.
The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log \rho_\phi(\f(\X))\)
and \(\nabla_\f \log p(\f(\X))\). \(\nabla_\f \log \rho_\phi(\f(\X))\) is likely 
intractable, considering \(\rho_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f(\X))\)
can be easy to compute analytically. To reduce variance in the gradients, we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log \rho_\phi(\f(\X))\), they use the Spectral Stein Gradient Estimator (SSGE)
\citep{shi_spectral_2018}. In short, this is a method for estimating the gradient
function of implicit distributions using approximations to eigenfunctions of a 
kernel-based operator. They bring three hyperparameters into the algorithm, the number of samples \(k\) from the implicit distribution used to approximate
the gradient, the number of eigenvectors \(J\) used to approximate the gradient, and
\(\eta\), a regularization parameter that smooths the gradient function.

@@

\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge. The training curves
of our model compared to that of standard NoisyNet-DQN on the four selected OpenAI
Gym~\citep{brockman_openai_2016} environments is shown in Figure~\ref{fig:training_curves}.
We hypothesize that the added KL-term will push the algorithm to retain weight uncertainty high
to keep a higher uncertainty in the output. While both methods can
find an optimal policy in Cartpole, our method uses considerably fewer frames
to do so.

@@

\cite{fortunato_noisy_2019} noted that the learned variance in their weight increased
in some games despite there existing an optimal deterministic solution, and the loss
provides no incentive to maintain any uncertainty. Figure~\ref{fig:mean_std} shows
the mean-absolute standard deviation for the penultimate and final layer. It is 
interesting to see that the standard deviation for NoisyNet continues to decrease
throughout the training while LFM decreases faster but then flattens out earlier.
This seems to indicate that it has found a stable policy, where optimizing further would
be overfitting to noise.

@@

\subsection{Informative Prior}
One of the benefits of a functional prior is that we can incorporate domain knowledge to
get more efficient exploration. We will now see how our method can utilize domain
knowledge to improve sample efficiency. To measure the effectiveness of priors we will use
a distribution that has a higher concentration of large Q-values for actions that we want
to incentivize.

\begin{align}
    p(\bm{s},a) &\sim \mathcal{N}(\bm{\mu}, \sigma I ), \text{ where}\\
    \bm{\mu} &= \lambda_1 + \lambda_2 \mathrm{A}_a \bm{s} \nonumber
\end{align}

% \begin{table}
%     \centering
%     \caption{Prior parameters}\label{tab:hyperparams}
%     \begin{tabular}{cccl}
%       \toprule % from booktabs package
%       Environment & \(\lambda_1\) & \(\lambda_2\) & \(\mathrm{A}\) \\
%       \midrule % from booktabs package
%       Cartpole & 0 & 100 & \(\begin{bmatrix} 0 & 0 & 1 & 1 \\
%                                              0 & 0 & -1 & -1
%                              \end{bmatrix}\)\\
%       MountainCar & 100 & 100 & \(\begin{bmatrix} 0 & -1 \\
%                                                   0 & 0 \\
%                                                   0 & 1
%                                 \end{bmatrix}\)\\
%       Acrobot & 100 & 100 & \(\begin{bmatrix} 0 & -1 & 0 & -1 & 0 & 0 \\
%                                                0 & 0 & 0 & 0 & 0 & 0 \\
%                                                0 & 1 & 0 & 1 & 0 & 0
%                             \end{bmatrix}\)\\
%       \bottomrule % from booktabs package
%     \end{tabular}
% \end{table}
